---
title: "An Introduction to R Markdown and Data Extraction Using R"
author: "Fadel Megahed"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: 
  html_document:
    code_download: TRUE
    code_folding: show
    number_sections: TRUE
    theme: simplex
    toc: TRUE
    toc_float: TRUE
  pdf_document:
    number_sections: TRUE
    toc: TRUE
bibliography: refs.bib 
csl: "american-journal-of-epidemiology.csl"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      verbose = FALSE,
                      cache = TRUE,
                      dpi = 300,
                      dev = c('png', 'postscript', 'tiff'),
                      out.width = '100%')
```

# Packages and Installation

```{r packages}
if(require(pacman) == FALSE) install.packages('pacman')
pacman::p_load(tidyverse, magrittr, rvest, lubridate)
```


# Extracting Data in R

## From a CSV File Online

### Apple's Mobility Data

1. Go to <https://covid19.apple.com/mobility>.  
2. Right Click on the "All Data CSV Button" and then click on "inspect".  
3. It will bring you close to the 'csv link' and you will need to find it by going up/down the inspector gadget (you will be close) and look for HTML code that contains <a href>.  Based on `r format(Sys.Date(), '%B %d, %Y')`, the link was <https://covid19-static.cdn-apple.com/covid19-mobility-data/2109HotfixDev13/v3/en-us/applemobilitytrends-2021-06-02.csv>.

```{r appleData}
appleMobility = read.csv('https://covid19-static.cdn-apple.com/covid19-mobility-data/2109HotfixDev13/v3/en-us/applemobilitytrends-2021-06-02.csv')

```



### GitHub
Once you have identified the dataset/repo, you are interested in. Click on 'Raw' and you will have the link you will need.  

1. For example, if you want to read [this file](https://github.com/owid/covid-19-data/blob/master/public/data/vaccinations/us_state_vaccinations.csv).  
2. Click on Raw and you will find the link to be updated to the link in the code below.

```{r vaccines}
vaccines = read.csv("https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/vaccinations/us_state_vaccinations.csv")
```



### Wikipedia Tables

```{r scrapeData}
wikiURL = 'https://en.wikipedia.org/w/index.php?title=COVID-19&oldid=1026813774'

wikiURL %>% read_html() %>% html_nodes("table.wikitable > tbody") %>% html_table() -> wikiTables

wikiTables[[1]]
```


# References {-}
<div id="refs"></div>


# Appendix

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.



